Surface-driven registration method for the structure-informed segmentation of diffusion MR images\tnoteref{t1}

### 2015-11-03 Publications/2015-NeuroImage/03-introduction.tex
modifications for MJLC comments

\revcomment[rephrased]{%
Here we present a segmentation and surface-to-volume registration method
  named \regseg{}, and show its usefulness in mapping anatomical information from structural
  into native \gls*{dmri} space to aid subsequent processing steps
  (reconstruction, tractography and network building using a cortical parcellation).
The underlying hypothesis is that the registration and segmentation problems
  in \gls*{dmri} can be solved simultaneously.}
\revcomment[C1 (R.1) \\ C10 (R.2)]{%
To implement \regseg{} we first establish an active-contours without edges
  \citep{chan_active_2001} segmentation framework.}
\revcomment[C2 (R.1)]{%
A specific set of reference surfaces extracted from the same subject initialize
  the 3D active-contours, which evolve searching for homogeneous regions in the multivariate
  target-image.
In the presented application to segment \gls*{dmri} data, we propose to use \regseg{}
  using a set of nested surfaces extracted from a strutrucal image (e.g. \gls*{t1}) and
  a bivariate target-volume comprehending the \gls*{fa} and \gls*{adc} maps as inputs.
The evolution of the surfaces is supported by a B-spline basis, optimized
  iteratively using a descent approach driven by shape-gradients
  \citep{besson_dream2s_2003,herbulot_segmentation_2006}.
Therefore, \regseg{} establishes a registration framework that actually
  overcomes the nonlinear warping induced by \gls*{epi} distortions.%
}
\Regseg{} integrates the benefits of segmentation and registration methods together and
  exploits the multivariate nature of \gls*{dmri} data to contribute in the proposed
  application on three key aspects:
  1) it avoids the resampling the \glspl*{dwi} introduced by the correction of \gls*{epi}
    distortions, since processing can be performed on the native \gls*{dmri} space;
  2) the surfaces are typically extracted from the \gls*{t1} image of the same subject,
    therefore \regseg{} does not require additional MR acquisitions to the minimal
    \gls*{dmri} protocol; and
  3) \regseg{} increases the geometrical accuracy of the overall process.
In this paper, we first verify the functionality of the method and the \regseg{}
  implementation using a set of digital phantoms, demonstrating the sub-pixel accuracy
  in registration.
We then evaluate \regseg{} on real \gls*{dmri} datasets, using a derivation of our
  instrumentation framework \citep{esteban_simulationbased_2014} which simulates
  known and realistic \gls*{epi} distortions.
We also compare \regseg{} and a nonlinear registration method to map the \lowb{} to
  the corresponding \gls*{t2} image of the same subject.
This method is a derivation of the \gls*{t2b} correction methods \citep{kybic_unwarping_2000},
  reproducing the settings of an existing implementation included within a widely used diffusion
  processing software (\emph{ExploreDTI}, \cite{leemans_exploredti_2009}).
With this comparison, we demonstrate how \regseg{} achieves higher accuracy with the
  simultaneous registration and segmentation process.

>>>>

\revcomment[rephrased]{%
Here we present a segmentation and surface-to-volume registration method
  named \regseg{}, and show its usefulness in mapping anatomical information from structural
  space into native \gls*{dmri} space to aid subsequent processing steps
  (reconstruction, tractography and network building using a cortical parcellation).
The underlying hypothesis is that the registration and segmentation problems
  in \gls*{dmri} can be solved simultaneously.}
\revcomment[C1 (R.1) \\ C10 (R.2)]{%
To implement \regseg{} we first establish an active-contours without edges
  \citep{chan_active_2001} segmentation framework.}
\revcomment[C2 (R.1)]{%
A specific set of reference surfaces extracted from the same subject initialize
  the 3D active-contours, which evolve searching for homogeneous regions in the multivariate
  target-image.
In the presented application to segment \gls*{dmri} data, we propose to use \regseg{}
  using a set of nested surfaces extracted from a strutrucal image (e.g. \gls*{t1}) and
  a bivariate target-volume comprehending the \gls*{fa} and \gls*{adc} maps as inputs.
The evolution of the surfaces is supported by a B-spline basis, optimized
  iteratively using a descent approach driven by shape-gradients
  \citep{besson_dream2s_2003,herbulot_segmentation_2006}.
Therefore, \regseg{} establishes a registration framework that actually
  deals with the nonlinear warping induced by \gls*{epi} distortions.%
}
\Regseg{} integrates the benefits of segmentation and registration methods together and
  exploits the multivariate nature of \gls*{dmri} data to contribute in the proposed
  application on three key aspects:
  1) it avoids the resampling the \glspl*{dwi} introduced by the correction of \gls*{epi}
    distortions, since processing can be performed on the native \gls*{dmri} space;
  2) the surfaces are typically extracted from the \gls*{t1} image of the same subject,
    therefore \regseg{} does not require additional MR acquisitions to the minimal
    \gls*{dmri} protocol; and
  3) \regseg{} increases the geometrical accuracy of the overall process.
In this paper, we first verify the functionality of the method and the \regseg{}
  implementation using a set of digital phantoms, demonstrating the sub-pixel accuracy
  in registration.
We then evaluate \regseg{} on real \gls*{dmri} datasets, using a derivation of our
  instrumentation framework \citep{esteban_simulationbased_2014} which simulates
  known and realistic \gls*{epi} distortions.
We also compare \regseg{} and a nonlinear registration method to map the \lowb{} to
  the corresponding \gls*{t2} image of the same subject.
This method is a derivation of the \gls*{t2b} correction methods \citep{kybic_unwarping_2000},
  reproducing the settings of an existing implementation included within a widely used diffusion
  processing software (\emph{ExploreDTI}, \cite{leemans_exploredti_2009}).
With this comparison, we demonstrate how \regseg{} achieves higher accuracy with the
  simultaneous registration and segmentation process.

### 2015-11-03 Publications/2015-NeuroImage/04-methods.tex
modifications for MJLC comments

The adaptation of this protocol to the simulated phantoms and real data is explained in the
  following sections.
\revcomment[C20 (R.2)]{%
A first proof of concept to demonstrate \regseg{} in phantoms with simple
  geometries is introduced, using $U_\text{true}$ without directional restrictions.
Then, a more realistic environment with $U_\text{true}$ restricted to the
  \gls*{pe} axis of real \gls*{dmri} datasets is presented.}

@@

\begin{figure*}
\includegraphics[width=\linewidth]{figure03}
\caption{Experimental workflow employed to process real data from the \acrfull*{hcp}.
  1) $\gammaset_R$ were extracted from the anatomical reference (\gls*{t1} image).
  2) For use as the ground truth, we generated a plausible synthetic distortion $U_\text{true}$
    from the field map with \eqref{eq:regseg-fieldmap}.
  3) The \gls*{dmri} data were warped using $U_\text{true}$ to reproduce the effects of real
    susceptibility-derived distortions.
  Target diffusion scalars (\gls*{fa} and \gls*{adc}) were computed with the distorted data and
    stacked to feed the multivariate input required by \regseg{}.
  4) The method was run to obtain $U_\text{test} = \hat{U}_\text{true}$, i.e., the estimate of
    the ground-truth deformation.
  5) The results were evaluated visually and quantitatively.
  \revcomment{%
  The arrows point to edges in the target images (yellow arrows for \gls*{fa}, blue for \gls*{adc}) that
    should be aligned with a surface, showing how distortion breaks the mapping with the structural space in which
    the contours are defined.}}\label{fig:regseg-evworkflows}
  \todo[inline]{C7 (R.1) \& C23 (R.2)}%
\end{figure*}

\paragraph*{Data}
To evaluate \regseg{} using real \gls*{dmri} data obtained from human brains,
  we collected 16 subjects from the \gls*{hcp} database.
The original acquisitions are released within ``unprocessed'' packages, whereas
  the ``minimally preprocessed'' packages contain the corresponding images after
  some processing (correction for several artifacts, brain-extraction, spatial
  normalization, etc.).
We refer the reader to \citep{essen_human_2012} for exact details of the acquisition
  parameters and \citep{glasser_minimal_2013} for the preprocessing issues.
These datasets comprise a large set of images, including \gls*{t1}, \gls*{t2}, and
  multi-shell \gls*{dmri} images.
\revcomment[C7 (R.1)]{%
Since we obtained the \gls*{dmri} data from the minimally preprocessed package, these
  images are corrected for \gls*{epi} distortions and spatially normalized in
  \gls*{t1} space.}

@@

\paragraph*{Settings} %
\revcomment[C5 (R.1)]{%
\Regseg{} accepts an affine mapping from surface-space to the \gls*{dmri} data as initialization.
However, the images provided by the \gls*{hcp} are already spatially normalized.
Therefore, the initial estimation of distortion is zero in this experiment.}
\revcomment[C20 (R.2)]{%
Since the distortion $U_\text{true}$ is aligned along the \gls*{pe} direction ($y$-axis in our settings),
  \regseg{} was configured to allow nonzero displacements only on that corresponding direction.}
\revcomment[C6 (R.1)]{%
For the experiments on real data, \regseg{} established a multi-resolution pyramid of
  B-Spline basis, with control points distributed on grids of the following spacings:
  40$\times$100$\times$40 [mm] for the first (coarser level), \isores[mm]{30} for the second,
  and 20$\times$30$\times$10 [mm] in the third.
Only the first and second levels included Gaussian smoothing of the target image ($\sigma$ = [2.0, 0.5] mm,
  respectively).
The actual choices of the parameter settings are publicly distributed with the source code for the
  experiments\footnote{\url{https://github.com/oesteban/RegSeg/tree/master/Scripts/pyacwereg/data/regseg_hcp.json}}.}
These settings were obtained manually, driven by the feedback obtained from the post-registration
  convergence reports \citepalias[like that found in][section S1.3]{esteban_useful_2015}.
We released \regseg{} along with the tool to generate such convergence reports.

\paragraph*{Cross-comparison} %
A dual workflow to the general evaluation framework used for \regseg{} (\autoref{fig:regseg-evworkflows}),
  was employed to integrate the alternate \gls*{t2b} registration scheme.
We reproduced the solution and settings provided with \emph{ExploreDTI}
  \citep{leemans_exploredti_2009}, which is a widely used toolkit for tractography analysis of
  \gls*{dti}.
\emph{ExploreDTI} internally employs \emph{elastix} \citep{klein_elastix_2010} to
  perform registration.
\revcomment[C21 (R.2)]{%
The deformation field is correspondingly restricted to the \gls*{pe} direction.
Settings used with \emph{elastix} are conveniently available in its format\footnote{\url{https://github.com/oesteban/RegSeg/blob/master/Scripts/pyacwereg/data/t2b_elastix_y.txt}}.}

>>>>

The adaptation of this protocol to the simulated phantoms and real data is explained in the
  following sections.
\revcomment[C20 (R.2)]{%
A first proof of concept is introduced to demonstrate \regseg{} in digital phantoms
  with simple geometries, using $U_\text{true}$ without directional restrictions.
Then, \regseg{} is evaluated in a framework using undistorted \gls*{dmri} datasets,
  and $U_\text{true}$ is derived from the corresponding inhomogeneity fieldmap of
  the subject.
Therefore, the deformation field is nonzero only in the \gls*{pe} axis,
  and resembles a real \gls*{epi} distortion.}

@@

\begin{figure*}
\includegraphics[width=\linewidth]{figure03}
\caption{Experimental workflow employed to process real data from the \acrfull*{hcp}.
  1) $\gammaset_R$ were extracted from the anatomical reference (\gls*{t1} image).
  2) For use as the ground truth, we generated a plausible synthetic distortion $U_\text{true}$
    from the field map with \eqref{eq:regseg-fieldmap}.
  3) The \gls*{dmri} data were warped using $U_\text{true}$ to reproduce the effects of real
    susceptibility-derived distortions.
  Target diffusion scalars (\gls*{fa} and \gls*{adc}) were computed with the distorted data and
    stacked to feed the multivariate input required by \regseg{}.
  4) The method was run to obtain $U_\text{test} = \hat{U}_\text{true}$, i.e., the estimate of
    the ground-truth deformation.
  5) The results were evaluated visually and quantitatively.
  \revcomment{%
  The arrows point to edges in the target images (light-yellow arrows for \gls*{fa}, blue for \gls*{adc}) that
    should be aligned with a surface, showing how distortion limits the direct mapping from the structural space
    in which the contours are defined.}}\label{fig:regseg-evworkflows}
  \todo[inline]{C7 (R.1) \& C23 (R.2)}%
\end{figure*}

\paragraph*{Data}
To evaluate \regseg{} using real \gls*{dmri} data obtained from human brains,
  we collected 16 subjects from the \gls*{hcp} database.
The original acquisitions are released within ``unprocessed'' packages, whereas
  the ``minimally preprocessed'' packages contain the corresponding images after
  some processing (correction for several artifacts, brain-extraction, spatial
  normalization, etc.).
We refer the reader to \citet{essen_human_2012} for exact details of the acquisition
  parameters and \citet{glasser_minimal_2013} for the preprocessing issues.
These datasets comprise a large set of images, including \gls*{t1}, \gls*{t2}, and
  multi-shell \gls*{dmri} images.
\revcomment[C7 (R.1)]{%
Since we obtained the \gls*{dmri} data from the minimally preprocessed package, these
  images are corrected for \gls*{epi} distortions and spatially normalized in
  \gls*{t1} space.}

@@

\paragraph*{Settings} %
\revcomment[C5 (R.1)]{%
\Regseg{} accepts an affine mapping from surface-space to the \gls*{dmri} data as initialization.
However, the images provided by the \gls*{hcp} are already spatially normalized.
Therefore, the initial estimation of distortion is zero in this experiment.}
\revcomment[C20 (R.2)]{%
Since the distortion $U_\text{true}$ is aligned along the \gls*{pe} direction ($y$-axis in our settings),
  \regseg{} was configured to allow nonzero displacements only on that corresponding direction.}
\revcomment[C6 (R.1)]{%
For the experiments on real data, \regseg{} established a multi-resolution pyramid of
  B-Spline functions, with control points distributed on grids of the following spacings:
  40$\times$100$\times$40 [mm] for the first (coarser) level, \isores[mm]{30} for the second level,
  and 20$\times$30$\times$10 [mm] in the third level.
Only the first and second levels included Gaussian smoothing of the target image ($\sigma$ = [2.0, 0.5] mm,
  respectively).
The actual choices of the parameter settings are publicly distributed with the source code for the
  experiments\footnote{\url{https://github.com/oesteban/RegSeg/tree/master/Scripts/pyacwereg/data/regseg_hcp.json}}.}
These settings were obtained manually, driven by the feedback obtained from the post-registration
  convergence reports \citepalias[like that found in][section S1.3]{esteban_useful_2015}.
We released \regseg{} along with the tool to generate such convergence reports.

\paragraph*{Cross-comparison} %
A dual workflow to the general evaluation framework used for \regseg{} (\autoref{fig:regseg-evworkflows}),
  was employed to integrate the alternate \gls*{t2b} registration scheme.
We reproduced the solution and settings provided with \emph{ExploreDTI}
  \citep{leemans_exploredti_2009}, which is a widely used toolkit for tractography analysis of
  \gls*{dti}.
\emph{ExploreDTI} internally employs \emph{elastix} \citep{klein_elastix_2010} to
  perform registration.
\revcomment[C21 (R.2)]{%
The deformation field is correspondingly restricted to the \gls*{pe} direction.
The settings file for \emph{elastix} is also
  available\footnote{\url{https://github.com/oesteban/RegSeg/blob/master/Scripts/pyacwereg/data/t2b_elastix_y.txt}}.}

### 2015-11-03 Publications/2015-NeuroImage/05-results.tex
update supplementals citation
modifications for MJLC comments

\begin{figure*}[!h]
  \centering
  \includestandalone[width=\linewidth]{figure05}
  \caption{A. Visual assessment of the results obtained with the low resolution sets:
    ``gyrus'' (top left), ``L'' (top right), ``ball'' (bottom left),
    and ``box'' at (bottom right).
  The contours recovered after registration are represented in yellow.
  \Regseg{} achieved high accuracy because it determined the almost exact locations of the registered
    contours with respect to their ground truth positions (shown in green).
  The partial volume effect makes segmentation of the sulci a challenging problem with voxel-wise
    clustering methods, but they were successfully segmented with \regseg{}.
  B. Quantitative evaluation of registration errors in terms of the average Hausdorff distances between
    surfaces at low (left) and high (right) resolutions, which demonstrate that the errors were
    consistently below the size of the voxels.
    }\label{fig:regseg-phantom}
\end{figure*}
\begin{figure*}[!h]
  \centering
  \includestandalone[width=\linewidth]{figure06}
  \caption{A. Example of a visual assessment report, which was generated automatically by the evaluation tool.
    Each view shows one component of the input image (in this case, the \gls*{fa} map), the ground-truth locations
    of the surfaces (green contours), and the resulting surfaces obtained with the test method (yellow contours).
  The first two rows show axial slices for \regseg{} and the \acrfull*{t2b} method, while the last two rows
    show the corresponding sagittal views.
  The coronal view is omitted because it was the least informative due to the directional property
    of the distortions.
  Specific regions where \regseg{} outperformed \gls*{t2b} are enlarged.
  B. Violin plots of the error distributions for each surface, which show the voxel size of the \gls*{dmri} images
    (1.25 mm), thereby supporting the improved results obtained by \regseg{} with the proposed settings.
  }\label{fig:regseg-results_real}
\end{figure*}

\subsection{Evaluation using real datasets and cross-comparison}\label{sec:regseg-results_hcp}
Finally, we compared the performance of \regseg{} with that of the standard \gls*{t2b}
  method.
Summary reports for visual assessment of the 16 cases are included in \suppl{section S5}.
In \autoref{fig:regseg-results_real}, box A, the visual report is shown for one subject.
We computed the \gls*{swindex} \eqref{eq:regseg-swindex} of every surface after registration
  using both the \regseg{} and \gls*{t2b} methods.
Finally, to compare the results, we performed Kruskal-Wallis H-tests
  (a non-parametric alternative to ANOVA) on the warping indices for the three surfaces of
  interest selected in \autoref{sec:regseg-experiments_evaluation}
  ($\Gamma_{VdGM}$, $\Gamma_{WM}$, $\Gamma_{pial}$).
All of the statistical tests showed that the error distributions obtained with \regseg{} and
  \gls*{t2b} were significantly different, and the violin plots in box B of
  \autoref{fig:regseg-results_real} demonstrate that the errors were always larger with \gls*{t2b}.
We also show the 95\% CIs of the \gls*{swindex} for these surfaces.
The aggregate CI for \regseg{} was 0.56--0.66 [mm], whereas the \gls*{t2b} method
  yielded an aggregate CI of 2.05--2.39 [mm].
The results of the statistical tests and the CIs are summarized in \autoref{tab:results_real}.

>>>>

\begin{figure*}[!h]
  \centering
  \includestandalone[width=\linewidth]{figure05}
  \caption{A. Visual assessment of the results obtained with the low resolution sets:
    ``gyrus'' (top left), ``L'' (top right), ``ball'' (bottom left),
    and ``box'' at (bottom right).
  The contours recovered after registration are represented in yellow.
  \Regseg{} achieved high accuracy because it determined the almost exact locations of the registered
    contours with respect to their ground truth positions (shown in green).
  The partial volume effect makes segmentation of the sulci a challenging problem with voxel-wise
    clustering methods, but they were successfully segmented with \regseg{}.
  B. Quantitative evaluation of registration errors in terms of the average Hausdorff distances between
    surfaces at low (left) and high (right) resolutions, which demonstrate that the errors were
    consistently below the size of the voxels.
    }\label{fig:regseg-phantom}
\end{figure*}
\begin{figure*}[!h]
  \centering
  \includestandalone[width=\linewidth]{figure06}
  \caption{A. Example of a visual assessment report, which was automatically generated by the evaluation tool.
    Each view shows one component of the input image (in this case, the \gls*{fa} map), the ground-truth locations
    of the surfaces (green contours), and the resulting surfaces obtained with the test method (yellow contours).
  The first two rows show axial slices for \regseg{} and the \acrfull*{t2b} method, while the last two rows
    show the corresponding sagittal views.
  The coronal view is omitted because it was the least informative due to the directional property
    of the distortions.
  Specific regions where \regseg{} outperformed \gls*{t2b} are enlarged.
  B. Violin plots of the error distributions for each surface, which show the voxel size of the \gls*{dmri} images
    (1.25 mm), thereby supporting the improved results obtained by \regseg{} with the proposed settings.
  }\label{fig:regseg-results_real}
\end{figure*}

\subsection{Evaluation using real datasets and cross-comparison}\label{sec:regseg-results_hcp}
Finally, we compared the performance of \regseg{} with that of the standard \gls*{t2b}
  method.
Summary reports for visual assessment of the 16 cases are included in \citepalias[section S5]{esteban_useful_2015}.
In \autoref{fig:regseg-results_real}, box A, the visual report is shown for one subject.
We computed the \gls*{swindex} \eqref{eq:regseg-swindex} of every surface after registration
  using both the \regseg{} and \gls*{t2b} methods.
Finally, to compare the results, we performed Kruskal-Wallis H-tests
  (a non-parametric alternative to ANOVA) on the warping indices for the three surfaces of
  interest selected in \autoref{sec:regseg-experiments_evaluation}
  ($\Gamma_{VdGM}$, $\Gamma_{WM}$, $\Gamma_{pial}$).
All of the statistical tests showed that the error distributions obtained with \regseg{} and
  \gls*{t2b} were significantly different, and the violin plots in box B of
  \autoref{fig:regseg-results_real} demonstrate that the errors were always larger with \gls*{t2b}.
We also show the 95\% CIs of the \gls*{swindex} for these surfaces.
The aggregate CI for \regseg{} was 0.56--0.66 [mm], whereas the \gls*{t2b} method
  yielded an aggregate CI of 2.05--2.39 [mm].
The results of the statistical tests and the CIs are summarized in \autoref{tab:results_real}.

### 2015-11-03 Publications/2015-NeuroImage/06-discussion.tex
modifications for MJLC comments

% -*- root: 00-main.tex -*-
\section{Discussion}
\label{sec:regseg-discussion}
\revcomment[C10 (R.2)]{%
We present \regseg{}, a simultaneous segmentation and registration method that
  maps a set of nested surfaces into a multivariate target-image.}
\revcomment[C1 (R.1)]{%
The nonlinear registration process evolves driven by the fitness of the
  picewise-smooth classification of voxels in the target volume imposed
  by the current mapping of the surfaces.}
\revcomment[moved]{%
We propose \regseg{} to map anatomical information extracted from \gls*{t1}
  images into the corresponding \gls*{dmri} of the same subject.
Previously, joint segmentation and registration has been applied successfully to
  other problems such as longitudinal object tracking \citep{paragios_level_2003}
  and atlas-based segmentation \citep{gorthi_active_2011}.
The most common approach involves optimizing a deformation model (registration)
  that supports the evolution of the active contours (segmentation),
  like \cite{paragios_level_2003,yezzi_variational_2003}.}
\revcomment[C1 (R.1)]{%
\Regseg{} can be seen as a particular case of atlas-based segmentation-registration methods,
  replacing the atlas by the structural image of the subject (\emph{structure-informed segmentation}).
In atlas-based segmentation, the origin of the problem is the anatomical variability hindering the
  atlas matching process.
Conversely, in structure-informed segmentation, the sources of variability are the geometrical distortions
  after imaging and the anatomical evolution in longitudinal studies.}
\revcomment[C3 (R.1)]{%
\cite{gorthi_active_2011} review the joint segmentation and registration methods, and propose a multiphase
  level set function derived from a labeled atlas for active contour-based registration in the atlas-based
  segmentation problem.
Alternatively, \regseg{} implements the active contours with a hierarchical set of explicit
  surfaces (triangular meshes) instead of the multiphase level sets, and registration
  is driven by shape-gradients \citep{herbulot_segmentation_2006}.
This solution is better adapted to the structure-informed segmentation, and it is possibly
   simpler to implement than the multiphase level set functions.}

\revcomment[C3 (R.1)]{%
An important antecedent of \regseg{} is \emph{bbregister} \citep{greve_accurate_2009}.
The tool has been widely adopted as the standard registration method to be used along with the \gls*{epi}
  correction of choice.
It implements a linear mapping and uses 3D active contours \emph{with edges} to
  search for intensity boundaries in the \lowb{} image.
The active contours are initialized using surfaces extracted from the \gls*{t1} using
  \emph{FreeSurfer} \citep{fischl_freesurfer_2012}.
To overcome the problem of nonlinear distortions, \emph{bbregister} excludes from the
  boundary search those regions that are typically warped.
Indeed, the distortion must be addressed separately because it is not supported by
  the affine transformation model.}
\revcomment[C22 (R.2)]{%
Conversely, the deformation model of \regseg{} is nonlinear and the active contours are
  \emph{without edges} \citep{chan_active_2001} because the \gls*{fa} and \gls*{adc} maps
  do not present steep image gradients (edges) but the anatomy can be identified
  by looking for piece-wise smooth homogeneous regions.}

@@

We evaluated \regseg{} in a real environment using the experimental framework presented
  in \autoref{fig:regseg-evworkflows}.
We processed 16 subjects from the \gls*{hcp} database using both \regseg{}
  and an in-house replication of the \acrfull*{t2b} method.
\Regseg{} obtained very high accuracy, with an aggregate 95\% \gls*{ci} of 0.56--0.66 [mm], which was
  below the pixel size of 1.25 mm.
The misregistration error that remained after \regseg{} was significantly lower ($p <$ 0.01) than the
  error corresponding to the \gls*{t2b} correction according to Kruskal-Wallis H-tests
  (\autoref{tab:results_real}).
Visual inspections of all the results \citepalias[section S5]{esteban_useful_2015} and the violin plots in
  \autoref{fig:regseg-results_real} confirmed that \regseg{} achieved higher accuracy
  than the \gls*{t2b} method in our settings.
We carefully configured the \gls*{t2b} method using the same algorithm and the
  same settings employed in a widely-used tool for \gls*{dmri} processing.
However, cross-comparison experiments are prone to the so-called \emph{instrumentation bias}
  \citep{tustison_instrumentation_2013}.
Therefore, these results did not prove that \regseg{} \emph{is better than} \gls*{t2b},
  but indicated that \regseg{} is a reliable option in this application field.
\revcomment[C5 (R.1)]{%
Finally, we also contributed proposing a piecewise-smooth segmentation model defined by
  a selection of nested surfaces to partition the multispectral space
  comprehending the \gls*{fa} and the \gls*{adc} maps and ultimately identify these
  structures in \gls*{dmri} space.
We also demonstrated the smoothness of the objective function on five of the real datasets
  \citepalias[figure S2]{esteban_useful_2015}, taking advantage of the directional
  restriction of possible distortions.
However, convergence of \regseg{} can be severely affected when surfaces are not dense enough.
Using the digital phantoms, we decimated the surfaces with large factors and
  effectively displaced the minimum of the shape-gradients from the minimum of the
  objective function.
Thus, the surfaces must be densely sampled to enable the use of the presented method.
}%

@@

\revcomment[C9 (R.2)]{%
Beyond the presented application on \gls*{dmri} data, \regseg{} can be indicated in situations
  where there are precise surfaces delineating the structure, a target multivariate
  image in which the surfaces must be fitted, and the mapping between the surfaces and
  the volume encodes a relevant physiological information, such as the normal/abnormal
  development or the macroscopic dynamics of organs and tissues.
For instance, \regseg{} may be applied in fields like neonatal brain image segmentation
  in longitudinal MRI studies of the early developmental patterns \citep{shi_neonatal_2010}.
In these studies, the surfaces obtained in a mature timepoint of the brain are retrospectively
  propagated to the initial timepoints, regardless the changes of the contrast and spatial
  development between them.
More generally, \regseg{} may also be applied to the personalized study of longitudinal alteration
  of the brain using multispectral images, for instance in the case of traumatic brain
  injury \citep{irimia_structural_2014} or in the development of brain tumors
  \citep{weizman_semiautomatic_2014}.}

>>>>

% -*- root: 00-main.tex -*-
\section{Discussion}
\label{sec:regseg-discussion}
\revcomment[C10 (R.2)]{%
We present \regseg{}, a simultaneous segmentation and registration method that
  maps a set of nested surfaces into a multivariate target-image.}
\revcomment[C1 (R.1)]{%
The nonlinear registration process evolves driven by the fitness of the
  picewise-smooth classification of voxels in the target volume imposed
  by the current mapping of the surfaces.}
\revcomment[moved]{%
We propose \regseg{} to map anatomical information extracted from \gls*{t1}
  images into the corresponding \gls*{dmri} of the same subject.
Previously, joint segmentation and registration has been applied successfully to
  other problems such as longitudinal object tracking \citep{paragios_level_2003}
  and atlas-based segmentation \citep{gorthi_active_2011}.
The most common approach involves optimizing a deformation model (registration)
  that supports the evolution of the active contours (segmentation),
  like \cite{paragios_level_2003,yezzi_variational_2003}.}
% Conversely, in structure-informed segmentation, the sources of variability are the geometrical distortions
%   after imaging and the anatomical evolution in longitudinal studies.}
\revcomment[C3 (R.1)]{%
\Regseg{} can be seen as a particular case of atlas-based segmentation-registration methods,
  replacing the atlas by the structural image of the subject (\emph{structure-informed segmentation}).
The main difference of atlas-based segmentation and the application at hand is the resolution of the
  target image.
Atlas-based segmentation is typically applied on high-resolution images.
A comprehensive review of joint segmentation and registration methods applied in atlas-based
  segmentation is found in \citep{gorthi_active_2011}.
They also propose a multiphase level-set function derived from a labeled atlas to implement the
  active contours that drive the atlas registration.
Alternatively, \regseg{} implements the active contours with a hierarchical set of explicit
  surfaces (triangular meshes) instead of the multiphase level sets, and registration
  is driven by shape-gradients \citep{herbulot_segmentation_2006}.
The use of explicit surfaces enables segmenting \gls*{dmri} images below pixel resolution.}

\revcomment[C3 (R.1)]{%
An important antecedent of \regseg{} is \emph{bbregister} \citep{greve_accurate_2009}.
The tool has been widely adopted as the standard registration method to be used along with the \gls*{epi}
  correction of choice.
It implements a linear mapping and uses 3D active contours \emph{with edges} to
  search for intensity boundaries in the \lowb{} image.
The active contours are initialized using surfaces extracted from the \gls*{t1} using
  \emph{FreeSurfer} \citep{fischl_freesurfer_2012}.
To overcome the problem of nonlinear distortions, \emph{bbregister} excludes from the
  boundary search those regions that are typically warped.
Indeed, the distortion must be addressed separately because it is not supported by
  the affine transformation model.}
\revcomment[C22 (R.2)]{%
Conversely, the deformation model of \regseg{} is nonlinear and the active contours are
  \emph{without edges} \citep{chan_active_2001} since the \gls*{fa} and \gls*{adc} maps
  do not present steep image gradients (edges) but the anatomy can be identified
  by looking for piece-wise smooth homogeneous regions.}

@@

We evaluated \regseg{} in a real environment using the experimental framework presented
  in \autoref{fig:regseg-evworkflows}.
We processed 16 subjects from the \gls*{hcp} database using both \regseg{}
  and an in-house replication of the \acrfull*{t2b} method.
\Regseg{} obtained very high accuracy, with an aggregate 95\% \gls*{ci} of 0.56--0.66 [mm], which was
  below the pixel size of 1.25 mm.
The misregistration error that remained after \regseg{} was significantly lower ($p <$ 0.01) than the
  error corresponding to the \gls*{t2b} correction according to Kruskal-Wallis H-tests
  (\autoref{tab:results_real}).
Visual inspections of all the results \citepalias[section S5]{esteban_useful_2015} and the violin plots in
  \autoref{fig:regseg-results_real} confirmed that \regseg{} achieved higher accuracy
  than the \gls*{t2b} method in our settings.
We carefully configured the \gls*{t2b} method using the same algorithm and the
  same settings employed in a widely-used tool for \gls*{dmri} processing.
However, cross-comparison experiments are prone to the so-called \emph{instrumentation bias}
  \citep{tustison_instrumentation_2013}.
Therefore, these results did not prove that \regseg{} \emph{is better than} \gls*{t2b},
  but indicated that \regseg{} is a reliable option in this application field.
\revcomment[C5 (R.1)]{%
Finally, we also contributed proposing a piecewise-smooth segmentation model defined by
  a selection of nested surfaces to partition the multispectral space
  comprehending the \gls*{fa} and the \gls*{adc} maps and ultimately identify these
  structures in \gls*{dmri} space.
We also demonstrated the smoothness of the objective function on five of the real datasets
  \citepalias[figure S2]{esteban_useful_2015}, taking advantage of the directional
  restriction of possible distortions.
However, convergence of \regseg{} can be severely affected when surfaces are not dense enough.
Using the digital phantoms, we decimated the surfaces by large factors.
Severely decimated surfaces introduced a bias that displaced the zero of the gradients from
  the minimum of the objective function.
Thus, the surfaces must be densely sampled to avoid the shape-gradients drive the algorithm
  in a wrong direction impeding convergence.
}%

@@

\revcomment[C9 (R.2)]{%
Beyond the presented application on \gls*{dmri} data, \regseg{} can be indicated in situations
  where there are precise surfaces delineating the structure, a target multivariate
  image in which the surfaces must be fitted, and the mapping between the surfaces and
  the volume encodes relevant physiological information, such as the normal/abnormal
  development or the macroscopic dynamics of organs and tissues.
For instance, \regseg{} may be applied in fields like neonatal brain image segmentation
  in longitudinal MRI studies of the early developmental patterns \citep{shi_neonatal_2010}.
In these studies, the surfaces obtained in a mature timepoint of the brain are retrospectively
  propagated to the initial timepoints, regardless the changes of the contrast and spatial
  development between them.
More generally, \regseg{} may also be applied to the personalized study of longitudinal alteration
  of the brain using multispectral images, for instance in the case of traumatic brain
  injury \citep{irimia_structural_2014} or in the monitorization of brain tumors
  \citep{weizman_semiautomatic_2014}.}
